---
title: "Nobel Prize Winners- Story"


output: 
    html_document:
        toc: yes
        theme: cosmo
        highlight: tango
        code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE,warning=FALSE)
```

# Introduction:
The Nobel Prize is perhaps the worlds most well known scientific award. 
Except for the honor, prestige and substantial prize money the recipient also gets a gold medal showing Alfred Nobel (1833 - 1896) who established the prize. Every year it's given to scientists and scholars in the categories chemistry, literature, physics, physiology or medicine, economics, and peace.

# Loading Libraries
```{r}
library(tidyverse)
library(ggthemes)
library(RColorBrewer)
library(kableExtra)
library(knitr)
library(ggrepel)
library(scales)
library(gridExtra)
library(tidytext)
library(wordcloud)
library(lubridate)
```

# Sourcing Data
```{r}
nobel<-read.csv("../input/archive.csv")
```

##  Getting Glimpse of data
```{r}
str(nobel)
head(nobel,3)
```

```{r}
percent <-function(col,tab=nobel){
    tab %>% 
    filter_(!is.na(col))%>%
    group_by_(col)%>%
    summarise(tot=n())%>%
    mutate(percent=round(tot/sum(tot)*100))%>%
    arrange(desc(tot))
}

ccount <-function(col,tab=nobel){
    tab %>% 
    filter(!is.na(col))%>%
    group_by_(col)%>%
    summarise(cnt=n())%>%
    arrange(desc(cnt))
}

nobel$Age <- nobel$Year - as.integer(substr(nobel$Birth.Date, 1,4))
```

# EDA  
## Gender  
 Getting the count of nobel prize winners by Male and Female .  
```{r fig.width=10,fig.height=6}
kable(nobel %>% group_by(Sex)%>%
    count())
   
g1<-ccount(col="Sex")%>%filter(!Sex=="")%>%ggplot(aes(x=Sex,y=cnt,fill=Sex)) + geom_bar(stat='identity',alpha=0.5) + theme_fivethirtyeight()+
   theme(axis.text.x=element_text(angle=45,vjust=0.5),legend.position='none',plot.title = element_text(size=12)) +scale_fill_manual(values=palette(rainbow(3)))+labs(title="Prize won by Male/Female")+geom_text_repel(aes(label=cnt))

g2<-nobel %>% group_by(Category,Sex)%>%filter(!Sex=="")%>% summarize(c=n())%>%mutate(percent=c/sum(c))%>%ggplot(aes(Category,percent,fill=Sex))+geom_bar(stat="identity",alpha=0.5,position="Dodge")+scale_y_continuous(labels = percent_format())+theme_fivethirtyeight()+scale_fill_manual(values=rainbow(5))+labs(title="Prize won by Category and Gender")

g1

g2
```    
Around 893 individual males received nobel prizes till 2016, where as female has got only 50 .  


## Category  
The Nobel Prize is widely regarded as the most prestigious award available in the fields of literature, medicine, physics, chemistry, economics and activism for peace. 
Economics category there was only 83 laureates, because economics field was established sonce `r min(nobel$Year[nobel$Category == 'Economics'])`. 
Medicine field got the highest number of laureates `r ccount(col="Category")$Category[which.max(ccount(col="Category")$cnt)] `.
```{r}

ccount(col="Category")%>%filter(!Category=="")%>%ggplot(aes(x=reorder(Category,-cnt),y=cnt,fill=Category)) + geom_bar(stat='identity',alpha=0.5) + theme_fivethirtyeight()+
   theme(axis.text.x=element_text(angle=45,vjust=0.5),legend.position='none',plot.title = element_text(size=12)) +scale_fill_manual(values=rainbow(n=10))+labs(title="Prize by Category")+geom_text_repel(aes(label=cnt))    
   
```

## Sharing Prizes  
Nobel Prizes can be shared to one or more receipients or organizations for their collaboration on same field/topic.  
Though the average number of laureates per prize increased substantially during the 20th century, a prize may not be shared among more than three people, 
although the Nobel Peace Prize can be awarded to organizations of more than three people  
```{r}

nobel %>% group_by(Year,Category)%>%summarise(ct=n())%>%ggplot(aes(x=ct))+geom_bar(fill=rainbow(n=1),alpha=0.5)+ theme_fivethirtyeight()+
   theme(axis.text.x=element_text(vjust=0.5),legend.position='none',plot.title = element_text(size=12)) +labs(title="Laureates per Prize")
```

## Multiple Nobel Laureates  
Lauraeates/organizations may receive nobel prize more than once, for their work in same or different field/topic.  
```{r}
kable(nobel %>% group_by(Full.Name)%>%summarize(Prize_Count=n())%>%filter(Prize_Count>1)%>%arrange(desc(Prize_Count))%>%head(5),caption="Laureates won Prize more than once")
```


# Age  

## Age vs Year of nobel Prize Win  
```{r fig.width=8,fig.height=6}
ggplot(nobel,aes(Year,Age,size=Age,col=cut(Age,5)))+geom_point(alpha=0.5)+geom_smooth(color="orange",se=FALSE)+theme_fivethirtyeight()+
   theme(axis.text.x=element_text(vjust=0.5),legend.position='none',plot.title = element_text(size=12)) +labs(title="Age and Year at which Nobel Prize was Won")+
  geom_text_repel(aes(label=ifelse(Age<=25 |Age>87,as.character(Full.Name),"")),size=5)

```

## Age Distribution  
The average age for receiving nobel prize seems to alwasys above 60 or near to that.For **Peace** category their exists exceptions , 2 outliers
exists which was below 25.
```{r fig.width=6,fig.height=5}
nobel %>% ggplot(aes(Category,Age,fill=Category))+geom_boxplot(alpha=0.5)+geom_jitter()+theme_fivethirtyeight()+scale_fill_manual(values=rainbow(6))+
   theme(axis.text.x=element_text(vjust=0.5),legend.position='none',plot.title = element_text(size=12)) +labs(title="Age Distribution by Category")

```

##  Age Trend for receiving nobel price per Category  
Always exists a upward trend of age  per year in 5 categories except **Peace**, there it was downward trend.  

```{r fig.width=10,fig.height=13}

nobel %>%ggplot(aes(Year,Age,col=Category))+geom_point()+facet_wrap(~Category)+geom_smooth()+scale_color_manual(values=rainbow(n=6))
```

## Female Laureates per Decade  
There was a deep imbalance between male and female laureates. Lets plot the proportion of female laurates per decade. 
In ** Literature, Peace and Medicine** category female proportion has increased a lot, from 1970's.  
```{r fig.width=7,fig.height=5}

prop_female_winners <- nobel %>% mutate(female_winner=ifelse(Sex=="Female",TRUE,FALSE),male_winner=ifelse(Sex=="Male",TRUE,FALSE),
                                        decade=floor(Year/10)*10)%>%
group_by(decade,Category)%>%summarize(fproportion=mean(female_winner,na.rm=TRUE),mproportion=mean(male_winner,na.rm=TRUE))
    


prop_female_winners %>%ggplot(aes(decade,fproportion,color=Category,group=Category))+geom_line()+geom_point()+scale_y_continuous(labels=scales::percent,limits=c(0.0,1.0),expand=c(0,0))+scale_x_continuous(breaks = seq(1900, 2020, 10))+theme_fivethirtyeight()
```

## Male Laureates Proportion per decade  
In Peace category, contribution of male was reducing from 1970 onwards, can see a downward trend. Literature Category also has downward trend.
```{r fig.width=7,fig.height=5}
prop_female_winners %>%ggplot(aes(decade,mproportion,color=Category,group=Category))+geom_line(size=1)+geom_point()+scale_y_continuous(labels=scales::percent,limits=c(0.0,1.0),expand=c(0,0))+scale_x_continuous(breaks = seq(1900, 2020, 10))+theme_fivethirtyeight()
```


### Youngest Nobel Laureates  
**Malala** becomes the youngest person ever to win the nobel Prize in 2014 at the age of 17 for Peace,before to that was **Sir William Lawrence Bragg** at the age of 25 for physics in 1915. 
Bragg and his father, Sir William Henry Bragg, won the award jointly for their work with X-ray spectra, X-ray diffraction and crystal structure.
```{r}

kable(nobel %>% select(Age,Full.Name,Year,Category)%>%arrange(Age)%>%head(5))
```

### Oldest Nobel Laureates  
90-year-old professor **Leonid Hurwicz** became the oldest person to ever win (one-third of the Prize in Economics); at 88, writer Doris Lessing became the oldest woman (Literature)
```{r}

kable(nobel %>% select(Age,Full.Name,Year,Category)%>%arrange(desc(Age))%>%head(5))

```

### First Female Laureates per Category  
```{r}
kable(nobel %>% filter(Sex=="Female")%>%group_by(Category)%>%select(Year,Full.Name,Category,Birth.Country)%>%top_n(1,desc(Year)))
```

# LifeSpan  
In dataset, death date of the naobel laureates was available, with that let's plot the distribution of their lifespan  

```{r}
nobel$Birth.Date<-as.Date(nobel$Birth.Date)
nobel$Death.Date<-as.Date(nobel$Death.Date)
nobel$lifespan<-year(nobel$Death.Date)-year(nobel$Birth.Date)
ggplot(nobel,aes(x=lifespan))+geom_histogram(binwidth = 2,fill="red4",alpha=0.5)+ theme_fivethirtyeight()+
   theme(plot.title = element_text(size=12)) +labs(title="Lifespan of Nobel Laureates")
```

## Distribution per Category  
Average age of the nobel laureates lies above 80 in all categories.  
```{r}
nobel %>% ggplot(aes(Category,lifespan,fill=Category))+geom_boxplot(alpha=0.5)+geom_jitter()+theme_fivethirtyeight()+scale_fill_manual(values=rainbow(6))+
   theme(axis.text.x=element_text(vjust=0.5),legend.position='none',plot.title = element_text(size=12)) +labs(title="Lifespan Distribution by Category")
```

## Awarded age vs Lifespan
As we saw in above plot, average age was 80, lets compare the age at which they got their award(nobel prize) and age at their death.
```{r}
nobel %>%select(Category,Awarded_age=Age,lifespan)%>%gather(key="s",value="value",2:3)%>%ggplot(aes(Category,value,fill=s))+geom_boxplot()+geom_jitter(size=0.4)+scale_fill_manual(values=rainbow(n=2))+theme_fivethirtyeight()+
   theme(plot.title = element_text(size=12)) +labs(title="Comparing Lifespan and Awarded Age by Category",fill="")
```


# Country
## Country Nobel Laureates - Top  
```{r}
ccount(col="Birth.Country")%>%filter(!Birth.Country=="" & cnt>4)%>%ggplot(aes(x=reorder(Birth.Country,cnt),y=cnt,fill=Birth.Country)) + geom_bar(stat='identity',alpha=0.5) + theme_fivethirtyeight()+
   theme(axis.text.x=element_text(angle=45,vjust=0.5),legend.position='none',plot.title = element_text(size=12)) +scale_fill_manual(values=rainbow(n=50))+labs(title="Prize by Country")+geom_text(aes(label=cnt))+coord_flip()
```

USA is the dominant country in receiving the prizes, lets get the proportion of it.

### 5.1 USA - Year wise proportion  
USA is the top country in getting nobel prizes, lets find the proportion of contribution per decade.  
```{r}
prop_usa_winners <- nobel %>% 
mutate(usa_born_winner=ifelse(Birth.Country=="United States of America",TRUE,FALSE),
       decade= floor(Year/10)*10)%>% group_by(decade)%>%summarize(proportion=mean(usa_born_winner==TRUE,na.rm=TRUE))
       
prop_usa_winners %>%ggplot(aes(decade,proportion))+geom_line(color="red")+geom_point(color="orange")+scale_y_continuous(labels=scales::percent,limits=c(0.0,1.0),expand=c(0,0))+theme_fivethirtyeight()
```

## Laureatures by Birth Country {.tabset .tabset-fade}  
### Nobel Laureates  for Peace  

Plotting the individual laureates per country who got nobel prize for ** PEACE **.  
```{r fig.width=10,fig.height=8}
nobel_peace <- nobel %>% 
    filter(( Category == 'Peace') & Sex != '')%>%
    filter(!is.na(Birth.Country)) %>%
    group_by(Birth.Country) %>%
    rename(region = Birth.Country) %>%
    summarise(value = n())
   
    kable(nobel_peace %>%arrange(desc(value))%>% head(10),caption =" Top Countries  who got nobel prize for Peace")
    
    
worldMap <- map_data('world')
allCountries <- data.frame(region=unique(worldMap$region), stringsAsFactors = F)

peace_country<- right_join(nobel_peace, allCountries)

peace_country$value[is.na(peace_country$value)] <- 0

#laureateBorn <- laureateBorn %>% arrange(region)

ggplot(peace_country, aes(map_id = region)) + 
    geom_map(aes(fill = log1p(value)), map = worldMap, color='black', size=0.15) + 
    expand_limits(x = worldMap$long, y = worldMap$lat) +
    theme_few()+
    theme(legend.position = "bottom",
         axis.ticks = element_blank(), 
         axis.title = element_blank(), 
         axis.text =  element_blank()) +
    scale_fill_gradient2(low="blue",mid = "purple",
                       high = "red" ,
                        name="Number of Laureates",
                        #breaks=Breaks,
                        labels=c('1', '2', '5', '10', '15', '25')) +
    guides(fill = guide_colorbar(barwidth = 10, barheight = .75)) + 
    ggtitle("Peace Laureates per Country of Birth")
```

### Nobel Laureates   for Medicine  
Plotting the individual laureates per country who got nobel prize for ** Medicine **.  
```{r fig.width=10,fig.height=8}
nobel_all<- nobel %>% 
    filter(Sex != '' & Category=="Medicine")%>%
    filter(!is.na(Birth.Country)) %>%
    group_by(Birth.Country) %>%
    rename(region = Birth.Country) %>%
    summarise(value = n())
nobel_all%>%arrange(desc(value))
worldMap <- map_data('world')
allCountries <- data.frame(region=unique(worldMap$region), stringsAsFactors = F)

country<- right_join(nobel_all, allCountries)

country$value[is.na(peace_country$value)] <- 0

#laureateBorn <- laureateBorn %>% arrange(region)

ggplot(country, aes(map_id = region)) + 
    geom_map(aes(fill = log1p(value)), map = worldMap, color='black', size=0.15) + 
    expand_limits(x = worldMap$long, y = worldMap$lat) +
    theme_few()+
    theme(legend.position = "bottom",
         axis.ticks = element_blank(), 
         axis.title = element_blank(), 
         axis.text =  element_blank()) +
    scale_fill_gradient(low="blue",
                       high = "red" ,
                        name="Number of Laureates",
                        #breaks=Breaks,
                        labels=c('1','10' , '20','50','70','80')
                       ) +
    guides(fill = guide_colorbar(barwidth = 10, barheight = .75)  )  + 
    ggtitle("Medicine - Nobel Laureates per Country of Birth")
```


# Prize Motivation  
Nobel Laureates has given their motivation , finding the most frequent words used by them in their motivation texts.  

```{r}
nobel$Motivation<-as.character(nobel$Motivation)
motive<-nobel %>%unnest_tokens(word,Motivation)%>%anti_join(stop_words,by="word")
motive%>%
  count(word, sort = TRUE) %>%
  top_n(10) %>%
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot() +
    geom_col(aes(word, n), fill = rainbow(n=10)) +
    theme(legend.position = "none", 
          plot.title = element_text(hjust = 0.5),
          panel.grid.major = element_blank()) +
    xlab("") + 
    ylab("Word Count") +
    ggtitle("Most Frequently Used Words by Nobel Laureates") +
    coord_flip()

word_counts <- motive %>%
  count(word, sort = TRUE) 

wordcloud(words = word_counts$word, freq = word_counts$n, min.freq = 1,
          max.words=300,  
          colors=rainbow(100))

```

## Bigrams

```{r}
motivation_bigrams <- nobel %>%
  unnest_tokens(bigram, Motivation, token = "ngrams", n = 2)

bigrams_separated <- motivation_bigrams %>%
  separate(bigram, c("word1", "word2"), sep = " ")

bigrams_filtered <- bigrams_separated %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word) 
  

bigram_united <- bigrams_filtered %>%
  filter(word1 != word2) %>%
    unite(bigram, word1, word2, sep = " ")
bigram_counts <- bigram_united %>% 
  count(bigram, sort = TRUE)
  head(bigram_counts) 
  

 bigram_counts %>% arrange(desc(n))%>% head(20)%>%ggplot(aes(x=reorder(bigram,n),y=n))+geom_bar(stat="identity",fill=rainbow(n=20))+theme_fivethirtyeight()+labs(title="Most Frequently Used bigrams by Nobel Laureates")+coord_flip()+theme(legend.position = "none", 
          plot.title = element_text(hjust = 0.5),
          panel.grid.major = element_blank()) +
    xlab("") + 
    ylab("Word Count") +
        coord_flip()
```

## Network of Bigrams
```{r}
library(igraph)
library(ggraph)
bigram_count <- bigrams_filtered %>% 
  count(word1, word2, sort = TRUE)

head(bigram_count)
bigram_graph <- bigram_count %>%  filter(n > 4) %>%  graph_from_data_frame()

bigram_graph

set.seed(2017)


 
 
 a <- grid::arrow(type = "closed", length = unit(.15, "inches"))

ggraph(bigram_graph, layout = "fr") +
  geom_edge_link(aes(edge_alpha = n), show.legend = FALSE,
                 arrow = a, end_cap = circle(.07, 'inches')) +
  geom_node_point(color = "lightblue", size = 5) +
  geom_node_text(aes(label = name), vjust = 1, hjust = 1) +
  theme_void()
```


# Nobel Prizes per Year by Category    
```{r fig.width=10,fig.height=8}
nobel %>%filter(!Sex=="")%>% group_by(Year,Sex)%>%summarize(c=n())%>%arrange(Year)%>%ggplot(aes(x=Year,c,col=Sex,group=Sex))+geom_line(size=1,alpha=0.5)+scale_x_continuous(breaks = seq(1900, 2020, 5))+theme_fivethirtyeight()+theme(axis.text.x = element_text(angle=90))+labs(title="Prizes per Year")+scale_color_manual(values=rainbow(n=2))

top_10<-ccount(col="Birth.Country")%>%filter(!Birth.Country=="")%>%head(5)


nobel %>%filter(Birth.Country %in% top_10$Birth.Country)%>% group_by(Year,Birth.Country)%>%summarize(c=n())%>%arrange(Year)%>%ggplot(aes(x=Year,c,col=Birth.Country,group=1))+geom_line(size=1,alpha=0.5)+scale_x_continuous(breaks = seq(1900, 2020, 5))+theme_fivethirtyeight()+theme(axis.text.x = element_text(angle=90),legend.position = "none")+labs(title="Top Countries -Prizes per Year")+facet_wrap(~Birth.Country)+scale_y_continuous(breaks=seq(0,10,2))
```

# Laureate Type
Nobel Prizes will be issued either to individual or to an organization. 
```{r}
nobel %>%select(Laureate.Type) %>% ccount(col="Laureate.Type")%>%ggplot(aes(x=Laureate.Type,y=cnt,fill=Laureate.Type)) + geom_bar(stat='identity',alpha=0.5) + theme_fivethirtyeight()+
   theme(axis.text.x=element_text(angle=45,vjust=0.5),legend.position='none',plot.title = element_text(size=12)) +scale_fill_manual(values=palette(rainbow(3)))+labs(title="Prize by Laureate Type")+geom_text_repel(aes(label=cnt))


```