---
title: "Story of Nobel Prize Winners"
output:
  html_document:
    df_print: paged
---

#What is Nobel Prize?
The Nobel Prize is perhaps the worlds most well known scientific award.
It is a set of annual international awards bestowed in several categories by Swedish and Norwegian institutions in recognition of academic, cultural, or scientific advances.
Except for the honor, prestige and substantial prize money the recipient also gets a gold medal showing Alfred Nobel (1833 - 1896) who established the prize. Every year it's given to scientists and scholars in the categories chemistry, literature, physics, physiology or medicine, economics, and peace.

# Loading the required Libraries
```{r}
library(tidyverse)
library(ggthemes)
library(RColorBrewer)
library(kableExtra)
library(knitr)
library(ggrepel)
library(scales)
library(gridExtra)
library(tidytext)
library(wordcloud)
library(lubridate)
library(igraph)
library(ggraph)
```

# Reading the Dataset
This is the dataset with a record for every individual or organization that was awarded the Nobel Prize(since 1901).
The Data was acquired from the Nobel Prize API.

```{r}
nobel<-read.csv("D:/groot/Summer/archive.csv")
```

##  Let's have a glance at data
```{r}
str(nobel)
head(nobel,3)
```

## Pre-defining functions
```{r}
percent <-function(col,tab=nobel){
    tab %>% 
    filter_(!is.na(col))%>%
    group_by_(col)%>%
    summarise(tot=n())%>%
    mutate(percent=round(tot/sum(tot)*100))%>%
    arrange(desc(tot))
}

ccount <-function(col,tab=nobel){
    tab %>% 
    filter(!is.na(col))%>%
    group_by_(col)%>%
    summarise(cnt=n())%>%
    arrange(desc(cnt))
}

nobel$Age <- nobel$Year - as.integer(substr(nobel$Birth.Date, 1,4))
```

# Exploratory Data Analysis

## Gender  
 Finding the count of nobel prize winners by their gender.

```{r}
kable(nobel %>% group_by(Sex)%>%
    count())
   
```

```{r}
g1<-ccount(col="Sex")%>%filter(!Sex=="")%>%ggplot(aes(x=Sex,y=cnt,fill=Sex)) + geom_bar(stat='identity',alpha=0.5) + theme_fivethirtyeight()+
   theme(axis.text.x=element_text(angle=45,vjust=0.5),legend.position='none',plot.title = element_text(size=12)) +scale_fill_manual(values=palette(rainbow(3)))+labs(title="Prize won by Male/Female")+geom_text_repel(aes(label=cnt))

g1
```


```{r}
g2<-nobel %>% group_by(Category,Sex)%>%filter(!Sex=="")%>% summarize(c=n())%>%mutate(percent=c/sum(c))%>%ggplot(aes(Category,percent,fill=Sex))+geom_bar(stat="identity",alpha=0.5,position="Dodge")+scale_y_continuous(labels = percent_format())+theme_fivethirtyeight()+scale_fill_manual(values=rainbow(5))+labs(title="Prize won by Category and Gender")

g2

```

observation: 893 individual males received nobel prizes(till 2016), where as 50 females .

## Category

This prestigious award is available in the fields of literature, medicine, physics, chemistry, economics and activism for peace.

```{r}
ccount(col="Category")%>%filter(!Category=="")%>%ggplot(aes(x=reorder(Category,-cnt),y=cnt,fill=Category)) + geom_bar(stat='identity',alpha=0.5) + theme_fivethirtyeight()+
   theme(axis.text.x=element_text(angle=45,vjust=0.5),legend.position='none',plot.title = element_text(size=12)) +scale_fill_manual(values=rainbow(n=10))+labs(title="Prize by Category")+geom_text_repel(aes(label=cnt)) 
```


## Sharing Prizes  
Nobel Prizes can be shared to one or more receipients/organizations for their collaboration on same field/topic.  
Note : a prize can not be shared among more than three people, 
although the Nobel Peace Prize can be awarded to organizations of more than three people.


```{r}
nobel %>% group_by(Year,Category)%>%summarise(ct=n())%>%ggplot(aes(x=ct))+geom_bar(fill=rainbow(n=1),alpha=0.5)+ theme_fivethirtyeight()+
   theme(axis.text.x=element_text(vjust=0.5),legend.position='none',plot.title = element_text(size=12)) +labs(title="Laureates per Prize")
```
## Multiple Nobel Laureates  
Lauraeates/organizations may receive nobel prize more than once, for their work in same or different field/topic.

```{r}
kable(nobel %>% group_by(Full.Name)%>%summarize(Prize_Count=n())%>%filter(Prize_Count>1)%>%arrange(desc(Prize_Count))%>%head(5),caption="Laureates won Prize more than once")
```


# Age  

Let's explore Age parameter to gain some useful insights from data.

## Age vs Year of nobel Prize

```{r}
ggplot(nobel,aes(Year,Age,size=Age,col=cut(Age,5)))+geom_point(alpha=0.5)+geom_smooth(color="orange",se=FALSE)+theme_fivethirtyeight()+
   theme(axis.text.x=element_text(vjust=0.5),legend.position='none',plot.title = element_text(size=12)) +labs(title="Age and Year at which Nobel Prize was Won")+
  geom_text_repel(aes(label=ifelse(Age<=25 |Age>87,as.character(Full.Name),"")),size=5)
```

Observation : Malala Yousafzai is the youngest to recieve Nobel Prize.

## Age Distribution  
```{r}
nobel %>% ggplot(aes(Category,Age,fill=Category))+geom_boxplot(alpha=0.5)+geom_jitter()+theme_fivethirtyeight()+scale_fill_manual(values=rainbow(6))+
   theme(axis.text.x=element_text(vjust=0.5),legend.position='none',plot.title = element_text(size=12)) +labs(title="Age Distribution by Category")
```

## Age Trend : Nobel price per Category

```{r}
nobel %>%ggplot(aes(Year,Age,col=Category))+geom_point()+facet_wrap(~Category)+geom_smooth()+scale_color_manual(values=rainbow(n=6))
```

observation: upward trend of age per year in all categories except Peace

## Female Laureates per Decade

```{r}
prop_female_winners <- nobel %>% mutate(female_winner=ifelse(Sex=="Female",TRUE,FALSE),male_winner=ifelse(Sex=="Male",TRUE,FALSE),
                                        decade=floor(Year/10)*10)%>%
group_by(decade,Category)%>%summarize(fproportion=mean(female_winner,na.rm=TRUE),mproportion=mean(male_winner,na.rm=TRUE))
    
prop_female_winners %>%ggplot(aes(decade,fproportion,color=Category,group=Category))+geom_line()+geom_point()+scale_y_continuous(labels=scales::percent,limits=c(0.0,1.0),expand=c(0,0))+scale_x_continuous(breaks = seq(1900, 2020, 10))+theme_fivethirtyeight()
```

## Male Laureates Proportion per decade

```{r}
prop_female_winners %>%ggplot(aes(decade,mproportion,color=Category,group=Category))+geom_line(size=1)+geom_point()+scale_y_continuous(labels=scales::percent,limits=c(0.0,1.0),expand=c(0,0))+scale_x_continuous(breaks = seq(1900, 2020, 10))+theme_fivethirtyeight()
```

observation: There was a deep imbalance between male and female laureates.

## Youngest Nobel Laureates

```{r}
kable(nobel %>% select(Age,Full.Name,Year,Category)%>%arrange(Age)%>%head(5))
```
Observation: Malala Yousafzai is the youngest person ever to win the nobel Prize.


## Eldest Nobel Laureates

```{r}
kable(nobel %>% select(Age,Full.Name,Year,Category)%>%arrange(desc(Age))%>%head(5))
```
Observation: 90-year-old professor Leonid Hurwicz is the oldest person ever to win a Nobel Prize.


## First Female Laureates per Category

```{r}
kable(nobel %>% filter(Sex=="Female")%>%group_by(Category)%>%select(Year,Full.Name,Category,Birth.Country)%>%top_n(1,desc(Year)))
```

# LifeSpan

Using death date of the Nobel laureates, let’s plot their lifespan distribution

```{r}
nobel$Birth.Date<-as.Date(nobel$Birth.Date)
nobel$Death.Date<-as.Date(nobel$Death.Date)
nobel$lifespan<-year(nobel$Death.Date)-year(nobel$Birth.Date)
ggplot(nobel,aes(x=lifespan))+geom_histogram(binwidth = 2,fill="red4",alpha=0.5)+ theme_fivethirtyeight()+
   theme(plot.title = element_text(size=12)) +labs(title="Lifespan of Nobel Laureates")
```

```{r}
ggplot(nobel,aes(x=lifespan,fill=Category))+geom_histogram(binwidth = 2,alpha=0.5)+ theme_fivethirtyeight()+
   theme(plot.title = element_text(size=12)) +labs(title="Lifespan of Nobel Laureates by Category")
```

```{r}
nobel %>% ggplot(aes(Category,lifespan,fill=Category))+geom_boxplot(alpha=0.5)+geom_jitter()+theme_fivethirtyeight()+scale_fill_manual(values=rainbow(6))+
   theme(axis.text.x=element_text(vjust=0.5),legend.position='none',plot.title = element_text(size=12)) +labs(title="Lifespan Distribution by Category")
```
Observation: Average age of the nobel laureates lies near/above 80 in all categories.


```{r}
nobel %>%select(Category,Awarded_age=Age,lifespan)%>%gather(key="s",value="value",2:3)%>%ggplot(aes(Category,value,fill=s))+geom_boxplot()+geom_jitter(size=0.4)+scale_fill_manual(values=rainbow(n=2))+theme_fivethirtyeight()+
   theme(plot.title = element_text(size=12)) +labs(title="Comparing Lifespan and Awarded Age by Category",fill="")
```

# Category

```{r}
nobel %>%filter(!Sex=="")%>% group_by(Year,Sex)%>%summarize(c=n())%>%arrange(Year)%>%ggplot(aes(x=Year,c,col=Sex,group=Sex))+geom_line(size=1,alpha=0.5)+scale_x_continuous(breaks = seq(1900, 2020, 5))+theme_fivethirtyeight()+theme(axis.text.x = element_text(angle=90))+labs(title="Prizes per Year")+scale_color_manual(values=rainbow(n=2))
```

```{r}
top_10<-ccount(col="Birth.Country")%>%filter(!Birth.Country=="")%>%head(5)


nobel %>%filter(Birth.Country %in% top_10$Birth.Country)%>% group_by(Year,Birth.Country)%>%summarize(c=n())%>%arrange(Year)%>%ggplot(aes(x=Year,c,col=Birth.Country,group=1))+geom_line(size=1,alpha=0.5)+scale_x_continuous(breaks = seq(1900, 2020, 5))+theme_fivethirtyeight()+theme(axis.text.x = element_text(angle=90),legend.position = "none")+labs(title="Top Countries -Prizes per Year")+facet_wrap(~Birth.Country)+scale_y_continuous(breaks=seq(0,10,2))
```

```{r}
nobel %>%select(Laureate.Type) %>% ccount(col="Laureate.Type")%>%ggplot(aes(x=Laureate.Type,y=cnt,fill=Laureate.Type)) + geom_bar(stat='identity',alpha=0.5) + theme_fivethirtyeight()+
   theme(axis.text.x=element_text(angle=45,vjust=0.5),legend.position='none',plot.title = element_text(size=12)) +scale_fill_manual(values=palette(rainbow(3)))+labs(title="Prize by Laureate Type")+geom_text_repel(aes(label=cnt))
```

# Country

```{r}
ccount(col="Birth.Country")%>%filter(!Birth.Country=="" & cnt>4)%>%ggplot(aes(x=reorder(Birth.Country,cnt),y=cnt,fill=Birth.Country)) + geom_bar(stat='identity',alpha=0.5) + theme_fivethirtyeight()+
   theme(axis.text.x=element_text(angle=45,vjust=0.5),legend.position='none',plot.title = element_text(size=12)) +scale_fill_manual(values=rainbow(n=50))+labs(title="Prize by Country")+geom_text(aes(label=cnt))+coord_flip()
```


## USA - Year wise analysis
```{r}
prop_usa_winners <- nobel %>% 
mutate(usa_born_winner=ifelse(Birth.Country=="United States of America",TRUE,FALSE),
       decade= floor(Year/10)*10)%>% group_by(decade)%>%summarize(proportion=mean(usa_born_winner==TRUE,na.rm=TRUE))
       
prop_usa_winners %>%ggplot(aes(decade,proportion))+geom_line(color="red")+geom_point(color="orange")+scale_y_continuous(labels=scales::percent,limits=c(0.0,1.0),expand=c(0,0))+theme_fivethirtyeight()+labs(title="USA - Year wise analysis")
```

## India - Year wise analysis
```{r}
prop_india_winners <- nobel %>% 
mutate(india_born_winner=ifelse(Birth.Country=="India",TRUE,FALSE), decade= floor(Year/10)*10)%>% group_by(decade)%>%summarize(proportion=mean(india_born_winner==TRUE,na.rm=TRUE))
       
prop_india_winners %>%ggplot(aes(decade,proportion))+geom_line(color="blue")+geom_point(color="orange")+scale_y_continuous(labels=scales::percent,limits=c(0.0,1.0),expand=c(0,0))+theme_fivethirtyeight()+labs(title="India - Year wise analysis")
```

## Birth Country

### laureates per country who got nobel prize for "PEACE".
```{r}
nobel_peace <- nobel %>% 
    filter(( Category == 'Peace') & Sex != '')%>%
    filter(!is.na(Birth.Country)) %>%
    group_by(Birth.Country) %>%
    rename(region = Birth.Country) %>%
    summarise(value = n())
   
    kable(nobel_peace %>%arrange(desc(value))%>% head(10),caption =" Top Countries  who got nobel prize for Peace")
```
```{r}
library(maps)
```

```{r}
worldMap <- map_data('world')
allCountries <- data.frame(region=unique(worldMap$region), stringsAsFactors = F)

peace_country<- right_join(nobel_peace, allCountries)

peace_country$value[is.na(peace_country$value)] <- 0

#laureateBorn <- laureateBorn %>% arrange(region)

ggplot(peace_country, aes(map_id = region)) + 
    geom_map(aes(fill = log1p(value)), map = worldMap, color='black', size=0.15) + 
    expand_limits(x = worldMap$long, y = worldMap$lat) +
    theme_few()+
    theme(legend.position = "bottom",
         axis.ticks = element_blank(), 
         axis.title = element_blank(), 
         axis.text =  element_blank()) +
    scale_fill_gradient2(low="blue",mid = "purple",
                       high = "red" ,
                        name="Number of Laureates",
                        #breaks=Breaks,
                        labels=c('1', '2', '5', '10', '15', '25')) +
    guides(fill = guide_colorbar(barwidth = 10, barheight = .75)) + 
    ggtitle("Peace Laureates per Country of Birth")
```

### laureates per country who got nobel prize for "Medicine".

```{r}
nobel_all<- nobel %>% 
    filter(Sex != '' & Category=="Medicine")%>%
    filter(!is.na(Birth.Country)) %>%
    group_by(Birth.Country) %>%
    rename(region = Birth.Country) %>%
    summarise(value = n())
nobel_all%>%arrange(desc(value))
```

```{r}
worldMap <- map_data('world')
allCountries <- data.frame(region=unique(worldMap$region), stringsAsFactors = F)

country<- right_join(nobel_all, allCountries)

country$value[is.na(peace_country$value)] <- 0

#laureateBorn <- laureateBorn %>% arrange(region)

ggplot(country, aes(map_id = region)) + 
    geom_map(aes(fill = log1p(value)), map = worldMap, color='black', size=0.15) + 
    expand_limits(x = worldMap$long, y = worldMap$lat) +
    theme_few()+
    theme(legend.position = "bottom",
         axis.ticks = element_blank(), 
         axis.title = element_blank(), 
         axis.text =  element_blank()) +
    scale_fill_gradient(low="blue",
                       high = "red" ,
                        name="Number of Laureates",
                        #breaks=Breaks,
                        labels=c('1','10' , '20','50','70','80')
                       ) +
    guides(fill = guide_colorbar(barwidth = 10, barheight = .75)  )  + 
    ggtitle("Medicine - Nobel Laureates per Country of Birth")
```

### laureates per country who got nobel prize for "Physics".

```{r}
nobel_all<- nobel %>% 
    filter(Sex != '' & Category=="Physics")%>%
    filter(!is.na(Birth.Country)) %>%
    group_by(Birth.Country) %>%
    rename(region = Birth.Country) %>%
    summarise(value = n())
nobel_all%>%arrange(desc(value))
```

```{r}
worldMap <- map_data('world')
allCountries <- data.frame(region=unique(worldMap$region), stringsAsFactors = F)

country<- right_join(nobel_all, allCountries)

country$value[is.na(peace_country$value)] <- 0

#laureateBorn <- laureateBorn %>% arrange(region)

ggplot(country, aes(map_id = region)) + 
    geom_map(aes(fill = log1p(value)), map = worldMap, color='black', size=0.15) + 
    expand_limits(x = worldMap$long, y = worldMap$lat) +
    theme_few()+
    theme(legend.position = "bottom",
         axis.ticks = element_blank(), 
         axis.title = element_blank(), 
         axis.text =  element_blank()) +
    scale_fill_gradient(low="yellow",
                       high = "green" ,
                        name="Number of Laureates",
                        #breaks=Breaks,
                        labels=c('1','10' , '20','50','70','80')
                       ) +
    guides(fill = guide_colorbar(barwidth = 10, barheight = .75)  )  + 
    ggtitle("Physics - Nobel Laureates per Country of Birth")
```

# Prize Motivation

Nobel Laureates have given their motivation, let's find the most frequent words used by them in their motivation texts.

```{r}
nobel$Motivation<-as.character(nobel$Motivation)
motive<-nobel %>%unnest_tokens(word,Motivation)%>%anti_join(stop_words,by="word")
motive%>%
  count(word, sort = TRUE) %>%
  top_n(10) %>%
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot() +
    geom_col(aes(word, n), fill = rainbow(n=10)) +
    theme(legend.position = "none", 
          plot.title = element_text(hjust = 0.5),
          panel.grid.major = element_blank()) +
    xlab("") + 
    ylab("Word Count") +
    ggtitle("Most Frequently Used Words by Nobel Laureates") +
    coord_flip()
```

## Word Cloud

```{r}
word_counts <- motive %>%
  count(word, sort = TRUE) 

wordcloud(words = word_counts$word, freq = word_counts$n, min.freq = 1,
          max.words=300,  
          colors=rainbow(100))
```

